Are environmental and geographic variables effective surrogates for genetic variation in conservation planning?

### 2016-04-29 code/rmarkdown/article.Rmd
+ changes to stats + surrogacy tests + figures, tables, SI

<!--  title page -->
Short running title: Genetic surrogates in conservation \newline
Word count: XXXXX \newline
Number of references: XXXX \newline
Number of figures, table and text boxes: 4 \newline
\newpage

@@

Intra-specific genetic variation is an important component for species persistence. As a consequence, there has been increasing interest in understanding how to design prioritizations that preserve and cultivate this variation [@r67; @r3; @r61]. Although the strength of natural selection is a continuous force; broadly speaking, genetic variation can be classified as either adaptive or neutral [reviewed in @r480]. Adaptive genetic variation is associated with loci that (significantly) affect fitness. Typically, such "adaptive" loci are detected due to anomalous patterns among individuals [eg. @r456; @r487; @r480; @r490; but see @r489 for alternative methods]. By preserving existing patterns of adaptive variation, protected areas can ensure that populations with particularly beneficial adaptations are not lost [@r3]. In contrast, neutral genetic variation is associated with loci that do not (significantly) affect fitness. Neutral variation often reflects the evolutionary history of different populations. By preserving existing patterns of neutral variation, protected areas can secure different genetic lineages and avoid the adverse effects of low genetic diversity [eg. inbreeding depression; @r67]. Thus protected areas networks should preserve both adaptive and neutral patterns of genetic variation for species of conservation interest [@r67; @r3]. Despite this, genetic data are not often used to inform prioritizations because they require considerable investment to obtain and analyze [@r61; but see @r5 and @r157 for developments on using genetic data in conservation planning]. Thus prioritizations may fail to capture species' intra-specific genetic variation, and ultimately fail to deliver long-term conservation outcomes.

@@

## Methods
### Study system
We used species distribution and genomic data collected by the IntraBioDiv project in the European Alps [@r451; see @r452 and @r478 for further explanation of data collection methods]. This data set contains genetic data for a diverse range of species that were collected using a standardized sampling regime. Additionally, the data set covers a large geographic domain that spans across a range of attitudinal gradients. As a consequence, there is large environmental variation among the grid cell cells which is desirable to assess effectiveness of environmental variables as surrogates for adaptive genetic variation. Thus this data set was particularly well suited for our study.

The IntroBioDiv project data set contains data for 27 alpine plant species that was collected using a 20' longitude by 21' latitude grid (approx. 22.3 km $\times$ 25 km; Figure S1). Project members visited every second grid cell, and if a species was detected in a cell, samples were collected from three individuals. They genotyped samples using amplified fragment length polymorphisms [AFLP; @r453], and constructed matrices denoting the presence/absence of polymorphisms at loci for each species (mean `r round(mean(sapply(spp.StructureData.LST, structurer:::n.loci)),2)` $\pm$ `r round(sd(sapply(spp.StructureData.LST, structurer:::n.loci)),1)` SD markers genotyped per species; see Table 1 in @r478 for the number markers for each species). Although AFLPs are not ideal as genomic representatives (cf. single nucleotide polymorphisms; SNPs), no multi-species genomic data set with similar properties currently exists. Thus the data set contains information describing the genomic properties of individuals for each species present in every second grid cell.

@@

### Landscape data
We examined the effectiveness of environmental and geographic surrogates for adaptive and neutral genetic variation (respectively). We used the grid cells used to sample genetic data as planning units to develop prioritizations. We used only planning units that contained at least a single individual ($n = `r nrow(grid.DF)`$) to reduce computational burden. To examine the potential effects of cost, we calculated the total human population density inside each planning unit [obtained at 1 km resolution from the Global Rural-Urban Mapping Project; \texttt{GRUMP V1}; @r472] and used this to represent acquisition cost.

@@

To describe the environmental characteristics of each grid cell, we used broad-scale climatic variables. We obtained data for 19 bioclimatic variables across the extent of the study area [approx. 1 km resolution; @r33]. We used this data set because it is commonly to map habitat suitability in conservation planning. These layers and the planning units were projected into an equal-area coordinate system ($\texttt{ESRI:102014}$). To reduce dimensionality, the 19 bioclimatic variables were clipped to the planning units and subject to a principal components analysis (PCA; using \texttt{ArcMap} version 10.2.2; Table S1). The first `r as.character(as.english(surrogate.params.LST[[MODE]]$number.components))` principal components (PCs) cumulatively explained `r round(pca.DF[[4]][surrogate.params.LST[[MODE]]$number.components],1)` % of the variation and were used to generate `r as.character(as.english(surrogate.params.LST[[MODE]]$number.components))` new layers. We calculated the average value for each PC layer in each planning unit, and used these values to characterize climatic variation among the planning units (Figure S2). Thus we constructed a three-dimensional environmental space--wherein each planning unit occupies a single point--as a potential surrogate for adaptive variation.

### Adaptive and neutral genetic data
To investigate the effectiveness of the surrogates for adaptive and neutral genetic variation, we first needed to identify which of the sampled loci--if any--were under selection. We used two outlier detection methods to achieve this. The basic premise underpinning such methods is that neutral loci are expected to exhibit a specific level of variation among populations, and loci that deviate from this expectation should be under selection [reviewed in @r480]. The advantage of these methods--in contrast with environmental association analyses [XXX]--is that they do not use environmental data to identify loci under selection, which would have biased our analysis. We omitted the species \textit{`r missing.species`} from all subsequent analysis because it was only associated with one genetic lineage and thus not compatible with one of the outlier detection methods. Loci identified as outliers in both analyses were treated as adaptive, and the remaining loci were treated as neutral.

First, we used multinominal-Dirichlet models implemented in \texttt{BayeScan} [version 2.1; @r456] to identify loci under selection. We adopted a similar methodology to Bothwell _et al._ [-@r455] and applied it to each of the species in the data set. To avoid falsely classifying loci as adaptive, we grouped individuals into genetic lineages for each species (Figures SXXX--SXXX). We used the number of genetic lineages in each species previously identified by Alvarez _et al._ [-@r478]. For each species, we used \texttt{Structure} [version 2.3.4; using same parameters as @r478 per run with `r structure.params.LST[[MODE]]$numruns` runs per species] and \texttt{ClumPP} [version 1.1.2; @r459] to group individuals into populations. We then ran $\texttt{Bayescan}$ for each species using these groupings [1:1 prior odds; `r bayescan.params.LST[[MODE]]$nbp` pilot runs; `r bayescan.params.LST[[MODE]]$burn` burn-in iterations; `r bayescan.params.LST[[MODE]]$n` post-burn-in iterations thinned by `r bayescan.params.LST[[MODE]]$thin` iterations; @r456]. To avoid false-positives, we omitted loci where the global frequency of the minor allele was $\geq 0.1$ (following guidelines in the user manual) and assumed they were neutral. Additionally, we omitted individuals from this analysis when their population membership was uncertain (maximum membership probability $< `r structure.params.LST[[MODE]]$probability.threshold`$). We ran `r bayescan.params.LST[[MODE]]$reps` replicates per species using a suitable false discovery rate [FDR $\leq `r bayescan.params.LST[[MODE]]$fdr`$; @r466].

Second, we used PCAs to identify outlier loci using methods implemented in the \texttt{pcadapt R} package [version 2.2; @r487]. For each of the 26 species, we ran a PCA and extracted the minimum number of components needed to secure `r pcadapt.params.LST[[MODE]]$min.variance.explained` % of the variation in loci. We then computed Mahalanobis distances [@r398] to detect outlier loci ($P < `r pcadapt.params.LST[[MODE]]$alpha.level`$). To enable comparisons between the two outlier detection analyses, we used the same individuals in this analysis as used in the $BayeScan$ analysis.

After classifying loci as adaptive or neutral, we mapped the main gradients of the adaptive (if detected) and neutral genetic variation for each species (Figures S6--S32). For each species, we discarded the population groupings, and partitioned adaptive and neutral loci into separate data sets. We used non-metric multi-dimensional scaling [NMDS; implemented in the \texttt{vegan R} pacakge; @r458] using Gower distances [@r464; using the \texttt{cluster R} package to accommodate sparsity; @r457] to derive continuous variables that described the main gradients of adaptive and neutral genetic variation for each species (Table S2). To ensure that the ordinations described a sufficient amount of the genomic variation, we ran successive scaling analyses with increasing dimensionality until this was achieved (maximum stress value $\leq$ `r nmds.params.LST[[MODE]]$max.stress`; `r nmds.params.LST[[MODE]]$trymax` random starts for each analysis). Since each grid cell had up to three samples per species, we used the average of the ordinated values associated with the individuals in each grid cell to express the typical genomic characteristics of individuals in the cells.

@@

### Prioritizations and statistical analyses
We used the unreliable representative and adequate prioritization (URAP) formulation of the reserve selection problem [@r465]. This formulation can identify the minimum number of planning units required to preserve both a proportion of the species' range (using amount-based targets) and a proportion of variation found across the species' range [using space-based targets and attribute spaces; _sensu_ @r465]. We treated the environmental and geographic surrogate variables as separate attribute spaces. We also treated the adaptive (if detected) and neutral variables as separate attribute spaces for each species. We set the demand points for each species and each attribute space using the values associated with the planning units. For every prioritization we generated, we calculated the proportion of the adaptive (if detected) and neutral genetic variation that every prioritization secured to assess its performance. Prioritizations associated with negative values--because they secured only a very small amount of genetic variation--were replaced with zeros to facilitate statistical analysis. We solved all reserve selection problems to within `r (gurobi.params.LST[[MODE]]$MIPGap*100)` % of optimality using the $\texttt{rapr}$ R package [@r465] and $\texttt{Gurobi}$ [@r463].

To address the first aim--to determine if the environmental and geographic variables correlate with adaptive and neutral genetic variation--we fit maximum likelihood population-effects (MLPE) models [implemented in the \texttt{ResistanceGA R} package; @r488]. We computed differences between the grid cells in terms of their environmental conditions and  geographic position. Additionally, for each species, we computed the differences between the grid cells in terms of the adaptive and neutral genetic characteristics of individuals that inhabited them (if detected). All differences were calculated using Euclidean distances. We then fit a set of MLPE models to the differences in the adaptive genetic characteristics among the grid cells and the differences in the environmental conditions for each species associated with adaptive genetic variation. We also fit a set of similar models using the differences in neutral genetic characteristics and the geographic distances for each species. For each model we fit an intercept-only null model, and tested surrogate variables whether they explained substantially more variation using $\chi^2$. We then applied Bonferroni corrections to each set of models (separately). Note that this analysis does not take into account spatial auto-correlation. The subsequent analyses explicitly control for this effect.

To address the second aim--to determine if the environmental and geographic variables are effect surrogates for adaptive and neutral genetic variation--we generated a collection of single-species prioritizations. For each species, we first generated two sets of prioritizations using increasing environmental and geographic-based targets (separately). Prioritizations were generated using the following surrogate-based targets: `r paste(rapr.params.LST[[MODE]]$surrogacy.analysis$surrogate.target,collapse=', ')`. We then create a set of null expectations with which to compare the surrogate-based prioritizations. These null expectations represented possible solutions obtained using conventional conservation planning methods that just use amount (eg. Marxan). For each species, for each target, we generated `r rapr.params.LST[[MODE]]$surrogacy.analysis$random.replicates` prioritizations by randomly selecting planning units occupied by the species. The number of randomly selected planning units was set to equal the number of planning units in the optimal prioritizations.

@@

## Results
<!-- Surrogacy analysis -->
The environmental and geographic-based variables were effective surrogates for adaptive and genetic variation (Figure 1). The prioritizations generated using greater surrogate-based targets secured more genetic variation (XXXX). In addition, prioritizations generated using environmental targets often secured more adaptive genetic variation than prioritizations generated using geographic-based targets did neutral genetic variation (XXXX). Only particularly high geographic targets ($\geq 0.95$ %) were able to secure a moderate proportion of neutral variation ($\geq XXXX$ %). Overall, there was a non-linear relationship between the surrogate-based targets used to generate the prioritizations and the proportion of genetic variation they secured.

<!-- Scenario analysis -->
To determine if the addition of environmental and geographic targets could improve the effectiveness of prioritizations in a more realistic context, we generated prioritizations according to three different scenarios (Figure 2). The prioritizations generated under the single-species scenario showed similar selection frequencies across the study area--regardless of the combination of targets used to generate them. On the other hand, the prioritizations generated under the multi-species scenario showed different selection frequencies depending on the targets used to generate them. The addition of surrogate- or genetic-based targets resulted in prioritizations that contained substantially more planning units. The prioritizations generated under the multi-species with cost scenario had different spatial distributions to the previous two scenarios. The prioritizations generated under this third scenario concentrated conservation effort towards areas with higher elevation.

>>>>

<!--  title page -->
Short running title: Genetic surrogates in conservation \newline
Word count: XXXXX \newline
Number of references: XXXX \newline
Number of figures: XXXX \newline
Number of tables: XXXX \newline
Number of text boxes: XXXX \newline
\newpage

@@

Intra-specific genetic variation is an important for long-term species persistence. As a consequence, there has been increasing interest in designing prioritizations that preserve and cultivate this variation [@r67; @r3; @r61]. Although the strength of natural selection is a continuous force; broadly speaking, genetic variation can be classified as either adaptive or neutral [reviewed in @r480]. Adaptive genetic variation is associated with loci that (significantly) affect fitness. Typically, such "adaptive" loci are detected due to anomalous patterns among individuals [eg. @r456; @r487; @r480; @r490; but see @r489 for alternative methods]. By preserving existing patterns of adaptive variation, protected areas can ensure that populations with particularly beneficial adaptations are not lost [@r3]. In contrast, neutral genetic variation is associated with loci that do not (significantly) affect fitness. Neutral variation often reflects the evolutionary history of different populations. By preserving existing patterns of neutral variation, protected areas can secure different genetic lineages and avoid the adverse effects of low genetic diversity [eg. inbreeding depression; @r67]. Thus protected area networks should preserve both adaptive and neutral patterns of genetic variation for species of conservation interest [@r67; @r3]. Despite this, genetic data are not often used to inform prioritizations because they require considerable investment to obtain and analyze [@r61; but see @r5 and @r157 for developments on using genetic data in conservation planning]. Thus prioritizations may fail to capture species' intra-specific genetic variation, and ultimately fail to deliver long-term conservation outcomes.

@@

## Methods
### Study system
We used species distribution and genomic data collected by the IntraBioDiv project in the European Alps [@r451; see @r452 and @r478 for further explanation of data collection methods]. This data set contains information for 27 alpine plant species (Figure 1a). It was collected using a 20' longitude by 21' latitude grid (approx. 22.3 km $\times$ 25 km; Figure S1). Project members visited every second grid cell, and if a species was detected in a cell, samples were collected from three individuals. They genotyped samples using amplified fragment length polymorphisms [AFLP; @r453], and constructed matrices denoting the presence/absence of polymorphisms at loci for each species (mean `r round(mean(sapply(spp.StructureData.LST, structurer:::n.loci)),2)` $\pm$ `r round(sd(sapply(spp.StructureData.LST, structurer:::n.loci)),1)` SD markers genotyped per species; see Table 1 in @r478 for the number markers for each species). Although AFLPs are not ideal as genomic representatives (cf. single nucleotide polymorphisms; SNPs), no multi-species genomic data set with similar properties currently exists. Thus the data set contains information describing the genomic properties of individuals for each species present in every second grid cell.

We chose this data set for several reasons. First, conservation planning exercises typically consider tens to hundreds of species with different evolutionary histories and life histories. To ensure that our results were relevant for conservation planners, we required a similar sized data set. This data set contains genetic data for a diverse range of plant species. Second, we required genetic data for all species to be collected under a standardized sampling scheme, and this data set used a grid to collect samples. Third, we required broad-scale environmental variation across the study area to test the effectiveness of environmental variables. This data set covers a large geographic domain that spans across attitudinal gradients. Thus this data set was particularly well suited for our study.

@@

### Landscape data
We examined the effectiveness of environmental and geographic surrogates for adaptive and neutral genetic variation (respectively). We used the grid cells used to collect samples by the IntraBioDiv project as planning units to develop prioritizations. We used only planning units that contained at least a single individual ($n = `r nrow(grid.DF)`$) to reduce computational burden. To examine the potential effects of cost, we calculated the total human population density inside each planning unit [obtained at 1 km resolution from the Global Rural-Urban Mapping Project; \texttt{GRUMP V1}; @r472] and used this to represent acquisition cost (Figure 1b).

@@

To describe the environmental characteristics of each grid cell, we used broad-scale climatic variables. We obtained data for 19 bioclimatic variables across the extent of the study area [approx. 1 km resolution; @r33]. We used this data set because it is commonly to map habitat suitability in conservation planning [r@XXXX]. These layers and the planning units were projected into an equal-area coordinate system ($\texttt{ESRI:102014}$). To reduce dimensionality, the 19 bioclimatic variables were clipped to the planning units and subject to a principal components analysis (PCA; using \texttt{ArcMap} version 10.2.2; Table S1). The first `r as.character(as.english(surrogate.params.LST[[MODE]]$number.components))` principal components (PCs) cumulatively explained `r round(pca.DF[[4]][surrogate.params.LST[[MODE]]$number.components],1)` % of the variation and were used to generate new layers. We calculated the average value for each PC layer in each planning unit, and used these values to characterize climatic variation among the planning units (Figure S2). Thus we constructed a three-dimensional environmental space--wherein each planning unit occupies a single point--as a potential surrogate for adaptive variation.

### Adaptive and neutral genetic data
To investigate the effectiveness of the surrogates for adaptive and neutral genetic variation, we first needed to identify which of the sampled loci--if any--were under selection. We used two outlier detection methods to achieve this. We omitted the species \textit{`r missing.species`} from all subsequent analyses because it was not compatible with one of the outlier detection methods (see below). The basic premise underpinning such methods is that neutral loci are expected to exhibit a specific level of variation among populations, and loci that deviate from this expectation should be under selection [reviewed in @r480]. The advantage of these methods--in contrast with environmental association analyses [@rXXX]--is that they do not use environmental data to identify loci under selection, which would have biased our analysis. Loci identified as outliers in both analyses were treated as adaptive and the rest were treated as neutral. To avoid false-positives, we omitted loci from both outlier detection methods where the global frequency of the minor allele was $\geq 0.1$ and assumed they were neutral.

First, we used multinominal-Dirichlet models implemented in \texttt{BayeScan} [version 2.1; @r456] to identify loci under selection. We adopted a similar methodology to Bothwell _et al._ [-@r455] and applied it to each of the species in the data set. To avoid falsely classifying loci as adaptive, we grouped individuals into genetic lineages for each species (Figures SXXX--SXXX). We used the number of genetic lineages in each species previously identified by Alvarez _et al._ [-@r478]. Since \textit{`r missing.species`} was previously identified as containing only one population, it was not compatible with \texttt{BayeScan}. For each species, we used \texttt{Structure} [version 2.3.4; using same parameters as @r478 per run with `r structure.params.LST[[MODE]]$numruns` runs per species] and \texttt{ClumPP} [version 1.1.2; @r459] to group individuals into populations. We then ran $\texttt{Bayescan}$ for each species using these groupings [1:1 prior odds; `r bayescan.params.LST[[MODE]]$nbp` pilot runs; `r bayescan.params.LST[[MODE]]$burn` burn-in iterations; `r bayescan.params.LST[[MODE]]$n` post-burn-in iterations thinned by `r bayescan.params.LST[[MODE]]$thin` iterations; @r456].  Additionally, we omitted individuals from this analysis when their population membership was uncertain (maximum membership probability $< `r structure.params.LST[[MODE]]$probability.threshold`$). We ran `r bayescan.params.LST[[MODE]]$reps` replicates per species using a suitable false discovery rate [q $\leq `r bayescan.params.LST[[MODE]]$fdr`$; @r466].

Second, we used PCAs to identify outlier loci using methods implemented in the \texttt{pcadapt R} package [version 2.2; @r487]. To enable comparisons between the two outlier detection analyses, we used the same individuals in this analysis as used in the $BayeScan$ analysis. For each of the 26 species, we ran a PCA and extracted the minimum number of components needed to secure `r pcadapt.params.LST[[MODE]]$min.variance.explained` % of the variation in loci. We then used Mahalanobis distances [@r398] to compute _P_ values and then q-values [using the \texttt{qvalue R} package; @r491]. Outlier loci were detected using a suitable FDR ($q < `r pcadapt.params.LST[[MODE]]$alpha.level`$).

After classifying loci as adaptive or neutral, we mapped the main gradients of the adaptive (if detected) and neutral genetic variation for each species (Figures S6--S32). For each species, we discarded the population groupings, and partitioned adaptive and neutral loci into separate matrices. We used non-metric multi-dimensional scaling [NMDS; implemented in the \texttt{vegan R} package; @r458] using Gower distances [@r464; using the \texttt{cluster R} package to accommodate sparsity; @r457] to derive continuous variables that described the main gradients of adaptive and neutral genetic variation separately for each species (Table S2). To ensure that the ordinations described a sufficient amount of the genomic variation, we ran successive scaling analyses with increasing dimensionality until this was achieved (maximum stress value $\leq$ `r nmds.params.LST[[MODE]]$max.stress`; `r nmds.params.LST[[MODE]]$trymax` random starts for each analysis). Since each grid cell had up to three samples per species, we used the average of the ordinated values associated with the individuals in each grid cell to express the typical genomic characteristics of individuals in the cells.

@@

## Prioritizations and statistical analyses
We used the unreliable representative and adequate prioritization (URAP) formulation of the reserve selection problem [@r465]. This formulation can identify the minimum number of planning units required to preserve both a proportion of the species' range (using amount-based targets) and a proportion of variation found across the species' range [using space-based targets and attribute spaces; _sensu_ @r465]. We treated the environmental and geographic surrogate variables as separate attribute spaces. We also treated the adaptive (if detected) and neutral variables as separate attribute spaces for each species. We set the demand points for each species and each attribute space using the values associated with the planning units. For every prioritization we generated, we calculated the proportion of the adaptive (if detected) and neutral genetic variation that every prioritization secured to assess its performance. Prioritizations associated with negative values--because they secured only a very small amount of genetic variation--were replaced with zeros to facilitate statistical analysis. We solved all reserve selection problems to within `r (gurobi.params.LST[[MODE]]$MIPGap*100)` % of optimality using the $\texttt{rapr}$ R package [@r465] and $\texttt{Gurobi}$ [@r463].

To address the second aim--to determine if the environmental and geographic variables are effective surrogates for adaptive and neutral genetic variation--we generated a collection of single-species prioritizations. For each species, we generated `r rapr.params.LST[[MODE]]$surrogacy.analysis$replicates` prioritizations by randomly selecting different combinations of planning units that were occupied by the species. We then computed the proportion of adaptive genetic variation and environmental variation secured in each prioritization. We then generated another set of random prioritizations for each species, and computed the proportion of neutral genetic variation and geographic variation secured in each of these prioritizations.

We then fit two sets of generalized linear model (GLM) with logit links. The first set of models were fit to the proportion of adaptive genetic variation secured by the prioritizations using the environmental variation secured secured in the prioritization for each species. Similarly, the second set of models assessed the relationship between the proportion of neutral genetic variation and the geographic variation secured in each prioritization for each species. We conducted likelihood ratio tests between each model and its corresponding null model, and applied Bonferroni corrections to assess significance. To assess the effectiveness of these variables as surrogates, we computed the Cragg and Uhler's pseudo-R$^2$ value for each model.

@@

## Results
<!-- summary of methods -->
We successfully mapped the main gradients of adaptive (if present) and neutral variation for each of the 26 species. The outlier detection methods detected adaptive genetic variation in `r length(env.surrogacy.LST)` species. Generally, the two outlier detection methods gave consistent results (mean `r round(mean(spp.outlier.methods.congruence, na.rm=TRUE),2)` % loci per species `r round(sd(spp.outlier.methods.congruence, na.rm=TRUE),2)` SD). Of the species that were associated with adaptive genetic variation, only a few loci were detected as under selection (mean `r round(mean(spp.outlier.loci.props, na.rm=TRUE),2)` % loci per species `r round(sd(spp.outlier.loci.props, na.rm=TRUE),2)` SD). After identifying the loci under selection, we used NMDS analyses to identify the main gradients of adaptive (if detected) and neutral variation for each species. Generally, only a couple of continuous dimensions were needed to sufficiently describe the adaptive genetic variation for each species (minimum $K$ `r min(adaptive.k)`; mean $K$ `r mean(adaptive.k)`; max $K$ `r max(adaptive.k)`; Table SXXX). On the other hand, many continuous dimensions were often needed to sufficiently describe the neutral genetic variation for each species (minimum $K$ `r min(neutral.k)`; mean $K$ `r mean(neutral.k)`; max $K$ `r max(neutral.k)`; Table SXXX).

<!-- surrogacy analysis -->
The surrogate variables correlated with the genetic characteristics of individuals found in the planning units for most species (Table SXXX). Specifically, environmental variation was found to significantly correlate with adaptive genetic variation for `r ` % of the species associated with loci under selection (_P_ < 0.05; mean `r ` $R^2$ $\pm$ `r ` SD). Additionally, geographic distances between planning units were found to significantly correlate with neutral genetic variation among planning units for `r ` % of the species (_P_ < 0.05; mean `r ` $R^2$ $\pm$ `r ` SD). These results suggest that prioritizations with different environmental characteristics tend to have individuals with different adaptive genetic characteristics. Similarly, planning units that are further apart tend to have individuals with different neutral genetic characteristics.

<!-- scenario analysis -->
To determine if the addition of environmental and geographic targets could improve the effectiveness of prioritizations in a more realistic context, we generated prioritizations according to three different scenarios (Figure 2). The prioritizations generated under the single-species scenario showed similar selection frequencies across the study area--regardless of the combination of targets used to generate them. On the other hand, the prioritizations generated under the multi-species scenario showed different selection frequencies depending on the targets used to generate them. The addition of surrogate- or genetic-based targets resulted in prioritizations that contained substantially more planning units. The prioritizations generated under the multi-species with cost scenario had different spatial distributions to the previous two scenarios. The prioritizations generated under this third scenario concentrated conservation effort towards areas with higher elevation.
